### 2021-03-18 appendix_chapter_3.Rmd
moving appendix for chapter 3 to seperate rmd

---
title: 'Chapter 3: Appendix'
output:
  word_document:
    reference_docx: markdown/frontiers_template.docx
  html_document:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
bibliography: references.bib
csl: markdown/frontiers.csl.txt
---
```{r software, message=FALSE, warning=FALSE, include=FALSE}
## Load software need for entire script

#library(computeMSD)
library(ggplot2)
library(rgdal)
library(gridExtra)
library(ggpubr)
library(ggord)
library(mapproj)
library(vegan)
library(dplyr)
library(ggpubr)
library(maps)
library(flextable)
library(officer)
library(tidync)
library(tidyverse)
library(hexbin)
library(lubridate)

```

# Appendix

## Domain features

```{r eval=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}

library(jpeg)
library(grid)
library(gridExtra)

vector_2 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day2.jpg")))
vector_5 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day5.jpg")))
vector_10 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day10.jpg")))
vector_15 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day15.jpg")))
vector_20 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day20.jpg")))
vector_25 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day25.jpg")))
vector_30 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day30.jpg")))

ggarrange(vector_2, vector_5, vector_10, vector_15, vector_20, vector_25, vector_30, labels = "AUTO")
```

## Additional comparisons

```{r additional sims, message=FALSE, echo=FALSE, warning=FALSE}

# Read in simulation data and data wrangling

## April

p_RK4_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
p_RK4_mean_April <- p_RK4_April %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

p_RK4_April$type <- "RK4_April" # add extra column 'type' to dataframe and assigned 'RK4' label

Min_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H100_W00_mean_April <- Min_Kfloat_H100_W00_April %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H100_W00_April$type <- "Min_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H100_W00_April$velocity <- with(Min_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))

Min_Kfloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H095_W05_mean_April <- Min_Kfloat_H095_W05_April %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H095_W05_April$type <- "Min_Kfloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H095_W05_April$velocity <- with(Min_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))

Min_Kfloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H090_W10_mean_April <- Min_Kfloat_H090_W10_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H090_W10_April$type <- "Min_Kfloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H090_W10_April$velocity <- with(Min_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))

Min_Kfloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H085_W15_mean_April <- Min_Kfloat_H085_W15_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H085_W15_April$type <- "Min_Kfloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H085_W15_April$velocity <- with(Min_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_Kfloat_H100_W00_mean_April <- Mean_Kfloat_H100_W00_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_Kfloat_H100_W00_April$type <- "Mean_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_Kfloat_H100_W00_April$velocity <- with(Mean_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_KFloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H095_W05_mean_April <- Mean_KFloat_H095_W05_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H095_W05_April$type <- "Mean_KFloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H095_W05_April$velocity <- with(Mean_KFloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_KFloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H090_W10_mean_April <- Mean_KFloat_H090_W10_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H090_W10_April$type <- "Mean_KFloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H090_W10_April$velocity <- with(Mean_KFloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_KFloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H085_W15_mean_April <- Mean_KFloat_H085_W15_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H085_W15_April$type <- "Mean_KFloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H085_W15_April$velocity <- with(Mean_KFloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H100_W00_mean_April <- Max_Kfloat_H100_W00_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H100_W00_April$type <- "Max_KFloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H100_W00_April$velocity <- with(Max_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H095_W05_mean_April <- Max_Kfloat_H095_W05_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H095_W05_April$type <- "Max_KFloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H095_W05_April$velocity <- with(Max_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H090_W10_mean_April <- Max_Kfloat_H090_W10_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H090_W10_April$type <- "Max_KFloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H090_W10_April$velocity <- with(Max_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H085_W15_mean_April <- Max_Kfloat_H085_W15_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H085_W15_April$type <- "Max_KFloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H085_W15_April$velocity <- with(Max_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))

total_data_April <- rbind(Min_Kfloat_H100_W00_April, Min_Kfloat_H095_W05_April, Min_Kfloat_H090_W10_April, Min_Kfloat_H085_W15_April, Mean_Kfloat_H100_W00_April, Mean_KFloat_H095_W05_April, Mean_KFloat_H090_W10_April, Mean_KFloat_H085_W15_April, Max_Kfloat_H100_W00_April, Max_Kfloat_H095_W05_April, Max_Kfloat_H090_W10_April, Max_Kfloat_H085_W15_April)

## July

p_RK4_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
p_RK4_mean_July <- p_RK4_July %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

p_RK4_July$type <- "RK4_July" # add extra column 'type' to dataframe and assigned 'RK4' label

Min_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H100_W00_mean_July <- Min_Kfloat_H100_W00_July %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H100_W00_July$type <- "Min_Kfloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H100_W00_July$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))

Min_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H095_W05_mean_July <- Min_Kfloat_H095_W05_July %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H095_W05_July$type <- "Min_Kfloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H095_W05_July$velocity <- with(Min_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))

Min_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H090_W10_mean_July <- Min_Kfloat_H090_W10_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H090_W10_July$type <- "Min_Kfloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H090_W10_July$velocity <- with(Min_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))

Min_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H085_W15_mean_July <- Min_Kfloat_H085_W15_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H085_W15_July$type <- "Min_Kfloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H085_W15_July$velocity <- with(Min_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_Kfloat_H100_W00_mean_July <- Mean_Kfloat_H100_W00_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_Kfloat_H100_W00_July$type <- "Mean_Kfloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_Kfloat_H100_W00_July$velocity <- with(Mean_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_KFloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H095_W05_mean_July <- Mean_KFloat_H095_W05_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H095_W05_July$type <- "Mean_KFloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H095_W05_July$velocity <- with(Mean_KFloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_KFloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H090_W10_mean_July <- Mean_KFloat_H090_W10_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H090_W10_July$type <- "Mean_KFloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H090_W10_July$velocity <- with(Mean_KFloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_KFloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H085_W15_mean_July <- Mean_KFloat_H085_W15_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H085_W15_July$type <- "Mean_KFloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H085_W15_July$velocity <- with(Mean_KFloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H100_W00_mean_July <- Max_Kfloat_H100_W00_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H100_W00_July$type <- "Max_KFloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H100_W00_July$velocity <- with(Max_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H095_W05_mean_July <- Max_Kfloat_H095_W05_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H095_W05_July$type <- "Max_KFloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H095_W05_July$velocity <- with(Max_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H090_W10_mean_July <- Max_Kfloat_H090_W10_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H090_W10_July$type <- "Max_KFloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H090_W10_July$velocity <- with(Max_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_July.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H085_W15_mean_July <- Max_Kfloat_H085_W15_July %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H085_W15_July$type <- "Max_KFloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H085_W15_July$velocity <- with(Max_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))

total_data_July <- rbind(Min_Kfloat_H100_W00_July, Min_Kfloat_H095_W05_July, Min_Kfloat_H090_W10_July, Min_Kfloat_H085_W15_July, Mean_Kfloat_H100_W00_July, Mean_KFloat_H095_W05_July, Mean_KFloat_H090_W10_July, Mean_KFloat_H085_W15_July, Max_Kfloat_H100_W00_July, Max_Kfloat_H095_W05_July, Max_Kfloat_H090_W10_July, Max_Kfloat_H085_W15_July)

## October

p_RK4_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
p_RK4_mean_Oct <- p_RK4_Oct %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label

Min_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H100_W00_mean_Oct <- Min_Kfloat_H100_W00_Oct %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Min_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H095_W05_mean_Oct <- Min_Kfloat_H095_W05_Oct %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H095_W05_Oct$type <- "Min_Kfloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Min_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H090_W10_mean_Oct <- Min_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H090_W10_Oct$type <- "Min_Kfloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Min_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H085_W15_mean_Oct <- Min_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H085_W15_Oct$type <- "Min_Kfloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_Kfloat_H100_W00_mean_Oct <- Mean_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_Kfloat_H100_W00_Oct$type <- "Mean_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_KFloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H095_W05_mean_Oct <- Mean_KFloat_H095_W05_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H095_W05_Oct$type <- "Mean_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_KFloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H090_W10_mean_Oct <- Mean_KFloat_H090_W10_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H090_W10_Oct$type <- "Mean_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Mean_KFloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H085_W15_mean_Oct <- Mean_KFloat_H085_W15_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H085_W15_Oct$type <- "Mean_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H100_W00_mean_Oct <- Max_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H100_W00_Oct$type <- "Max_KFloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H095_W05_mean_Oct <- Max_Kfloat_H095_W05_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H095_W05_Oct$type <- "Max_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H090_W10_mean_Oct <- Max_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H090_W10_Oct$type <- "Max_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

Max_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_Oct.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H085_W15_mean_Oct <- Max_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H085_W15_Oct$type <- "Max_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))

total_data_Oct <- rbind(Min_Kfloat_H100_W00_Oct, Min_Kfloat_H095_W05_Oct, Min_Kfloat_H090_W10_Oct, Min_Kfloat_H085_W15_Oct, Mean_Kfloat_H100_W00_Oct, Mean_KFloat_H095_W05_Oct, Mean_KFloat_H090_W10_Oct, Mean_KFloat_H085_W15_Oct, Max_Kfloat_H100_W00_Oct, Max_Kfloat_H095_W05_Oct, Max_Kfloat_H090_W10_Oct, Max_Kfloat_H085_W15_Oct)

```

### Comparison of mean trajectories

```{r, message=FALSE, echo=FALSE, warning=FALSE}
waterdrag_plot_April <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = Mean_Kfloat_H100_W00_mean_April, aes(x = lon, y = lat), color = 'blue')+
  geom_path(data = Mean_Kfloat_H100_W00_mean_April, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Max_Kfloat_H100_W00_mean_April, aes(x = lon, y = lat), color = 'red')+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

wwdrag_plot_April <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat), color = 'grey')+
  geom_path(data = Min_Kfloat_H090_W10_mean_April, aes(x = lon, y = lat), color = 'green')+
  geom_path(data = Min_Kfloat_H085_W15_mean_April, aes(x = lon, y = lat), color = 'pink')+
  geom_path(data = Mean_KFloat_H095_W05_mean_April, aes(x = lon, y = lat), color = 'yellow')+
  geom_path(data = Mean_KFloat_H090_W10_mean_April, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat), color = 'red')+
  geom_path(data = Max_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat), color = 'blue')+
  geom_path(data = Max_Kfloat_H090_W10_mean_April, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Max_Kfloat_H085_W15_mean_April, aes(x = lon, y = lat), color = 'red')+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

waterdrag_plot_July <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = Mean_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'blue')+
  geom_path(data = Mean_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Max_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'red')+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

wwdrag_plot_July <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = Min_Kfloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'grey')+
  geom_path(data = Min_Kfloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'green')+
  geom_path(data = Min_Kfloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'pink')+
  geom_path(data = Mean_KFloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'yellow')+
  geom_path(data = Mean_KFloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'red')+
  geom_path(data = Max_Kfloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'blue')+
  geom_path(data = Max_Kfloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Max_Kfloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'red')+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

waterdrag_plot_Oct <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Max_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'red')+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

wwdrag_plot_Oct <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = Min_Kfloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'grey')+
  geom_path(data = Min_Kfloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'green')+
  geom_path(data = Min_Kfloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'pink')+
  geom_path(data = Mean_KFloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'yellow')+
  geom_path(data = Mean_KFloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Mean_KFloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'red')+
  geom_path(data = Max_Kfloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
  geom_path(data = Max_Kfloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Max_Kfloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'red')+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=8, fig.cap="Figure 7: Comparison of mean trajectories. Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag. In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10. In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13. The reader is referred to tables 1 and 2 for details pertaining to each scenario."}

ggarrange(waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, labels = "AUTO", ncol = 2, nrow = 3)

```
In the April simulations, both passive and kelp particles followed the general north-westward flow of the region with clear entrainment of kelp particles in sub-mesoscale features. Kelp particles follow the north

### Comparison of density

```{r density plots additional, message=FALSE, warning=FALSE, echo=FALSE}
## Heatmaps of study domain, use previous data for study site map and histograms## Coastline of Southern Africa
load("map/south_africa_coast_hi_res.Rdata")

## April

heatmap_RK4_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = p_RK4_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("17") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H100_W00_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H100W100") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H095_W05_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  #ggtitle("H095W05") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H090_W10_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H090W10") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H085_W15_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H085W15") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_Kfloat_H100_W00_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
  geom_path(data = Mean_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H095_W05_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H090_W10_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H085_W15_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H100_W00_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H100_W00_mean_April ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H095_W05_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H090_W10_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H085_W15_April <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

## July

heatmap_RK4_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("30") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H100_W00_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H100W100") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H095_W05_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  #ggtitle("H095W05") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H090_W10_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H090W10") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H085_W15_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H085W15") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_Kfloat_H100_W00_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
  geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H095_W05_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H090_W10_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H085_W15_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H100_W00_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H100_W00_mean_July ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H095_W05_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H090_W10_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H085_W15_July <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

## October

heatmap_RK4_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("43") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H100_W00_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H100W100") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H095_W05_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  #ggtitle("H095W05") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H090_W10_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H090W10") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H085_W15_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H085W15") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_Kfloat_H100_W00_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H095_W05_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H090_W10 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H085_W15_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H100_W00_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H095_W05_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H090_W10_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Max_KFloat_H085_W15_Oct <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Max_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

## April

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time for scenarios 17 - 29, for April 2018 (winter). The plots are regular hexagons within the density of particles is calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}

figure_a_April <- ggarrange(heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("none"))

figure_b_April <- ggarrange(heatmap_RK4_April, legend = c("right"))

ggarrange(figure_a_April, figure_b_April, ncol = 1, nrow = 2)

```
## July

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time for scenarios 30 - 42, for July 2018 (winter). The plots are regular hexagons within the density of particles is calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}

figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"))

figure_b_July <- ggarrange(heatmap_RK4_July, legend = c("right"))

ggarrange(figure_a_July, figure_b_July, ncol = 1, nrow = 2)

```

## October

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 10: Comparison of density plots at the end run time for scenarios 43 - 55, for October 2018 (spring). The plots are regular hexagons within the density of particles is calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}

figure_a_Oct <- ggarrange(heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"))

figure_b_Oct <- ggarrange(heatmap_RK4_Oct, legend = c("right"))

ggarrange(figure_a_Oct, figure_b_Oct, ncol = 1, nrow = 2)

```

## Comparison of distances

```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE}

pos_April <- c("RK4_April",             ## April
         "Min_Kfloat_H100_W00_April",
         "Min_Kfloat_H095_W05_April",
         "Min_Kfloat_H090_W10_April",
         "Min_Kfloat_H085_W15_April",
         "Mean_Kfloat_H100_W00_April",
         "Mean_Kfloat_H095_W05_April",
         "Mean_Kfloat_H090_W10_April",
         "Mean_Kfloat_H085_W15_April",
         "Max_Kfloat_H100_W00_April",
         "Max_Kfloat_H095_W05_April",
         "Max_Kfloat_H090_W10_April",
         "Max_Kfloat_H085_W15_April")

dis_plot_April <- ggplot(data = total_data_April, aes(x = type, y = distance, show.legend = FALSE))+
  geom_boxplot(data = p_RK4_April, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
  #scale_x_discrete(limits = pos)+
  stat_compare_means(method = "kruskal.test", label = "p.signif")+
  theme_bw()+
  ylab("Distance (km)")+
  xlab("Scenario")+
  #coord_flip() +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

custom_labs_April <- c("17","18","19","20","21","22","23","24","25","26","27","28", "29") #create custom labels and add them to dis_plot

dis_plot_April <- dis_plot_April +
  scale_x_discrete(labels= custom_labs_April)

pos_July <- c("RK4_July",             ## July
         "Min_Kfloat_H100_W00_July",
         "Min_Kfloat_H095_W05_July",
         "Min_Kfloat_H090_W10_July",
         "Min_Kfloat_H085_W15_July",
         "Mean_Kfloat_H100_W00_July",
         "Mean_Kfloat_H095_W05_July",
         "Mean_Kfloat_H090_W10_July",
         "Mean_Kfloat_H085_W15_July",
         "Max_Kfloat_H100_W00_July",
         "Max_Kfloat_H095_W05_July",
         "Max_Kfloat_H090_W10_July",
         "Max_Kfloat_H085_W15_July")

dis_plot_July <- ggplot(data = total_data_July, aes(x = type, y = distance, show.legend = FALSE))+
  geom_boxplot(data = p_RK4_July, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
  #scale_x_discrete(limits = pos)+
  stat_compare_means(method = "kruskal.test", label = "p.signif")+
  theme_bw()+
  ylab("Distance (km)")+
  xlab("Scenario")+
  #coord_flip() +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

custom_labs_July <- c("30","31","32","33","34","35","36","37","38","39","40","41", "42") #create custom labels and add them to dis_plot

dis_plot_July <- dis_plot_July +
  scale_x_discrete(labels= custom_labs_July)

pos_Oct <- c("RK4_Oct",              ## Oct
         "Min_Kfloat_H100_W00_Oct",
         "Min_Kfloat_H095_W05_Oct",
         "Min_Kfloat_H090_W10_Oct",
         "Min_Kfloat_H085_W15_Oct",
         "Mean_Kfloat_H100_W00_Oct",
         "Mean_Kfloat_H095_W05_Oct",
         "Mean_Kfloat_H090_W10_Oct",
         "Mean_Kfloat_H085_W15_Oct",
         "Max_Kfloat_H100_W00_Oct",
         "Max_Kfloat_H095_W05_Oct",
         "Max_Kfloat_H090_W10_Oct",
         "Max_Kfloat_H085_W15_Oct")

dis_plot_Oct <- ggplot(data = total_data_Oct, aes(x = type, y = distance, show.legend = FALSE))+
  geom_boxplot(data = p_RK4_Oct, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
  #scale_x_discrete(limits = pos)+
  stat_compare_means(method = "kruskal.test", label = "p.signif")+
  theme_bw()+
  ylab("Distance (km)")+
  xlab("Scenario")+
  #coord_flip() +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

custom_labs_Oct <- c("43","44","45","46","47","48","49","50","51","52","53","54","55") #create custom labels and add them to dis_plot

dis_plot_Oct <- dis_plot_Oct +
  scale_x_discrete(labels= custom_labs_Oct)

```

```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8, fig.width=6, fig.cap="Figure 11: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for April (Autumn), July (winter), and October (spring). The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. The reader is referred to tables 1 and 2 for details pertaining to each scenario."}

ggarrange(dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 3)

```